Contextuality and the Born rule in fuzzy quantum theories

### 2017-04-08 prop_letter.tex
prop_letter.tex \paragraph{Quantum IVPM and Real-valued Probability Measure} Updating...

Recall when providing infinity measurement resource, the quantum probability
measures are simply induced by the Born rule for $d\ge3$. If we increase
the measurement resource, i.e., pick finer intervals, it is natural
to ask whether quantum IVPMs approach the Born rule. In order to understand
this question, it is instructive to consider the 2-dimensional non-Born
quantum probability measure~$\mu_{0}:\events\rightarrow[0,1]$ defined
as follow:
\begin{equation}
\mu_{0}(P)=\begin{cases}
\mu_{\ket{\ps}}^{B}(P) & \textrm{, if }P\in\events_{0}\textrm{ ;}\\
\mu_{\ket{0}}^{B}(P) & \textrm{, otherwis,}
\end{cases}
\end{equation}
where $\events_{0}=\left\{ \mathbb{0},\proj{\ps},\proj{\ms},\mathbb{1}\right\} $
is a sub-event space. In another word, $\mu_{0}$ is consistent with
the Born rule on every proper sub-event space, but not on the set
of all projectors.

Given a quantum IVPM~$\bar{\mu}$, the preceding idea motivates the
definition of \emph{core} relative to a sub-event space~$\events'\subseteq\events$
as follow.
\begin{equation}
\mathrm{core}\left(\bar{\mu},\events'\right)=\set{\rho}{\forall P\in\events'.~\mu_{\rho}^{B}\left(P\right)\in\bar{\mu}\left(P\right)}\textrm{ .}
\end{equation}
Although the elements of a core are density matrices, we may just
use the state vectors for pure states if it is clear according to
the context. For example, $\mathrm{core}\left(\iota_{\infty}\circ\mu_{0},\events_{0}\right)=\left\{ \ket{\ps}\right\} $
and $\mathrm{core}\left(\iota_{\infty}\circ\mu_{0},\events\right)=\emptyset$,
where $\circ$ denote function composition.

@@

\begin{example}[Three-dimensional quantum three-interval-valued probability
measure]\label{ex:three-dimensional-three-value} Given a three dimensional
Hilbert space with an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $.
Let $\mathscr{I}=\left\{ \necess,\imposs,\unknown\right\} $, $\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$,
$\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$, and
\begin{eqnarray}
 &  & \begin{split}\imposs & =\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})\\
 & =\bar{\mu}(\proj{\ps})=\bar{\mu}(\proj{\ps'})\textrm{ ,}
\end{split}
\\
 &  & \begin{split}\necess & =\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})\\
 & =\bar{\mu}(\mathbb{1}-\proj{\ps})=\bar{\mu}(\mathbb{1}-\proj{\ps'})\textrm{ ,}
\end{split}
\\
 &  & \unknown=\bar{\mu}(P)\qquad\textrm{otherwise.}
\end{eqnarray}
It is straightforward to verify that $\bar{\mu}$ is a quantum interval-valued
probability measure.

Now assume there exists a real-valued quantum probability measure
$\mu$ such that $\mu(P)\in\bar{\mu}(P)$ for every event (projection)~$P$.
We derive a contradiction as follows. Suppose $\mu(P)\in\bar{\mu}(P)$
then we must have $\mu(\proj{0})=\mu(\proj{\ps})=\mu(\proj{\ps'})=0$.
By Gleason's theorem, there is a mixed state~$\rho=\sum_{j=1}^{N}q_{j}\proj{\phi_{j}}$
such that $\mu\left(P\right)=\sum_{j=1}^{N}q_{j}\ip{\phi_{j}}{P\phi_{j}}$,
where $\sum_{j=1}^{N}q_{j}=1$ and $q_{j}>0$. However, no pure state~$\ket{\phi}$
can satisfy $\ip{\phi}{0}=\ip{\phi}{\ps}=\ip{\phi}{\ps'}=0$; a contradiction.
\end{example}

>>>>

Recall when providing infinity measurement resource, the quantum probability
measures are simply induced by the Born rule for $d\ge3$. If we increase
the measurement resource, i.e., pick finer intervals, it is natural
to ask whether quantum IVPMs approach the Born rule. In order to understand
this question, it is instructive to consider the 2-dimensional non-Born
quantum probability measure~$\mu_{0}:\events\rightarrow[0,1]$ defined
as follow:
\begin{equation}
\mu_{0}(P)=\begin{cases}
\mu_{\ket{\ps}}^{\mathrm{B}}(P) & \textrm{, if }P\in\events_{0}\textrm{ ;}\\
\mu_{\ket{0}}^{\mathrm{B}}(P) & \textrm{, otherwis,}
\end{cases}
\end{equation}
where $\events_{0}=\left\{ \mathbb{0},\proj{\ps},\proj{\ms},\mathbb{1}\right\} $
is a sub-event space. In another word, $\mu_{0}$ is consistent with
the Born rule on every proper sub-event space, but not on the set
of all projectors.

Given a quantum IVPM~$\bar{\mu}$, the preceding idea motivates the
definition of \emph{core} relative to a sub-event space~$\events'\subseteq\events$
as follow.\footnote{\yutsung{{I changed the definition of core} to density matrix. While
this change might make it easier to explain the core of quantum IVPMs,
it could increase the difficulty to explain classical IVPMs because
we might need to explain the meaning of density matrix and the Born
rule for classical probability measures...}}
\begin{equation}
\mathrm{core}\left(\bar{\mu},\events'\right)=\set{\rho}{\forall P\in\events'.~\mu_{\rho}^{\mathrm{B}}\left(P\right)\in\bar{\mu}\left(P\right)}\textrm{ .}
\end{equation}
Although the elements of a core are density matrices, we may just
use the state vectors for pure states if it is clear according to
the context. For example, $\mathrm{core}\left(\iota_{\infty}\circ\mu_{0},\events_{0}\right)=\left\{ \ket{\ps}\right\} $
and $\mathrm{core}\left(\iota_{\infty}\circ\mu_{0},\events\right)=\emptyset$,
where $\circ$ denote function composition.

@@

\begin{example}[Three-dimensional quantum three-interval-valued probability
measure]\label{ex:three-dimensional-three-value} Given a three dimensional
Hilbert space with an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $.
Let $\mathscr{I}=\left\{ \necess,\imposs,\unknown\right\} $, $\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$,
$\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$, and
\begin{eqnarray}
 &  & \begin{aligned}\imposs & =\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})\\
 & =\bar{\mu}(\proj{\ps})=\bar{\mu}(\proj{\ps'})\textrm{ ,}
\end{aligned}
\\
 &  & \begin{aligned}\necess & =\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})\\
 & =\bar{\mu}(\mathbb{1}-\proj{\ps})=\bar{\mu}(\mathbb{1}-\proj{\ps'})\textrm{ ,}
\end{aligned}
\\
 &  & \unknown=\bar{\mu}(P)\qquad\textrm{otherwise.}
\end{eqnarray}
It is straightforward to verify that $\bar{\mu}$ is a quantum interval-valued
probability measure.

Let $\events_{1}$ be the minimal event space containing both $\proj{0}$
and $\proj{\ps}$. Because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs$,
for any mixed state~$\rho\in\mathrm{core}\left(\bar{\mu},\events_{1}\right)$,
we have $\mu_{\rho}^{\mathrm{B}}(\proj{0})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps})=0$.
Hence, $\rho$ is actually the pure state orthogonal to both $\ket{0}$
and $\ket{\ps}$. The only choice of $\rho$ is $\ket{2}$, and it
is easy to verify $\ket{2}$ is actually in $\mathrm{core}\left(\bar{\mu},\events_{1}\right)$.

Similarly, let $\events_{2}$ be the minimal event space containing
both $\proj{0}$ and $\proj{\ps'}$. We have $\mathrm{core}\left(\bar{\mu},\events_{2}\right)=\left\{ \ket{1}\right\} $.
Therefore, 
\begin{equation}
\mathrm{core}\left(\bar{\mu},\events\right)\subseteq\mathrm{core}\left(\bar{\mu},\events_{1}\right)\cap\mathrm{core}\left(\bar{\mu},\events_{2}\right)=\emptyset\textrm{ ,}
\end{equation}
and must be empty.\end{example}

We found this kind of incompatibility can be reduced if we increase
the measurement resource, that is, using finer intervals. We conjecture
by consuming more and more measurement resource, it would be more
and more likely that there is an empty core relative to all projectors~$\events$.
